PTFO 8-8695: Two Stars, Two Signals, No Planet

### 2020-06-14 paper_review/response_20200605.txt

******************************************
REFEREE COMMENT
----------

Introduction: I do not think this is youngest hot Jupiter. That
should be true just for transiting systems, and it was the
youngest at discovery, but the two RV planets from ESPaDOnS are
younger (Donati et al. 2016 and Yu et al. 2017).  Both are hot
Jupiters and in Taurus. I believe the age of Tap 26 is more
ambiguous due to age spreads (multiple groups?) in Taurus, but
the age of V830 Tau is certainly <3 Myr.

RESPONSE
----------

> Thank you for this correction. We have qualified the relevant
> sentence as the youngest known *transiting* hot Jupiter, and
> added a footnote to cite the V830 Tau and TAP 26 results.

@@

RESPONSE
----------
> We have changed the wording of the last sentence of the first
> paragraph in the introduction to use the word
> "precess", rather than "torque".

******************************************
REFEREE COMMENT
----------

Section 2: The first paragraph discusses the full frame images,
while the next section discusses the 2m data. Why did the authors
use the FFIs when they have the (higher quality) 2m data?

RESPONSE
----------
> For the essential analysis of the study (i.e., extracting and
> fitting the long and short signals), we used the 2m SPOC
> lightcurve.
>
> Our discussion of the FFI lightcurves in Section 2 was motivated
> by the discontinuity in the light curve at BJD 2458488.3,
> which was eventually excluded from further analysis. We wanted
> to compare the FFI and 2m light curves to see whether this jump
> was an instrumental systematic or a software systematic.
>
> A secondary reason to introduce the concept of FFIs in Section
> 2 is that they are used to measure the variability amplitudes
> of the other members of 25 Ori-1 in Section 4.3.

@@

RESPONSE
----------
> We have updated the text to include this point -- thank you for
> making it.
> 
> We have also updated the text to reflect our hesitation about
> whether or not the phenomenon is associated with binarity,
> primarily because multiple photometric periods do not
> necessarily imply binarity.

******************************************
REFEREE COMMENT
----------

Section 5.4: The authors conclude this subsection with a
reasonable argument that this isn't a protoplanet. Does this same
argument not rule out gas/debris at the Keplerian coronation
radius? What could maintain such clouds for years/decades and
then have them jump phase but at the same period? It's hard to do
this if it is not somehow physically attached to the star in some
what (e.g., spots can do this).

RESPONSE
----------
> This comment caused us to consider more carefully the merits
> and weaknesses of the two hypotheses we discussed. 
> We ended up with a stronger preference for
> clumpy torus model over the spot occultation model. The reasons
> for the preference are given in the text: 1) the spot
> occultation model seems more geometrically contrived than we
> initially thought, and 2) the Halpha excess seen by
> Johns-Krull+16 (and also in unpublished observations by
> Stauffer of scallop shells) seems likely to require hydrogen in
> orbit near the corotation radius.
>
> In response to the question of: "why can the shifting phase be
> explained more easily by dust/gas clumps than the
> protoplanet?", we now reference D'Angelo and Spruit 2012 (see
> in particular their Section 1), which discusses the possibility
> of cyclic accretion, with periods of "stalled" mass
> accumulation outside the corotation radius alternating with
> accreting phases.

### 2020-06-14 paper/ms.tex
paper & report cleanup

The explanations that \replaced{are not yet}{have not yet been} ruled
out include {\it (i)} transiting clumps of \added{dust or }gas at the
Keplerian co-rotation radius; {\it (ii)} transits of enshrouded
protoplanets; {\it (iii)} occultations of \replaced{starspots}{hot and
cold spots on the stellar surface} by an optically thick disk.  The
first and last explanations have added appeal because they are
flexible enough to explain not only the transient and persistent-dip
M-dwarfs, but also the ``scallop shell'' M-dwarfs
\citep{stauffer_orbiting_2017}.  \replaced{Despite this appeal, the
possibility of distinct mechanisms explaining these distinct
variability classes remains open.}{ In the spot occultation scenario
\citep{zhan_complex_2019},
the ``blobs'' that provide the necessary
brightness contrast are on the surface of the star, rather than at the co-rotation
radius. The observing geometry would
require an inner ring to be projected into a narrow band in front of the
stellar disk, in order to produce variations with sufficiently
short timescales. 
% The outer disk would need to either not be present,
% or to be misaligned enough, and narrow enough, that it did not fall
% along the line of sight.
% JNW Doesn't this same problem affect the "dusty torus" model?
This geometric requirement seems too contrived to occur in even
$\lesssim$1\% of young M dwarf systems. }

\replaced{The evolution of \ptfo\ over}{Observations of \ptfo\ from}
the past decade could offer additional hints.  \deleted{Specifically,
}\ptfo's transition between having none, one, and multiple dips per
cycle seems important \added{(Figure~\ref{fig:o_minus_c})}.
%It
%strains the ``enshrouded protoplanet'' interpretation, because there
%are no known processes that cause a planet's orbital phase to jump.
% JNW Doesn't seem worth beating on a dead horse.
\deleted{The dips would then need to be caused by material that was
somehow disrupted from the planet, but somehow remained co-orbital for
an extended duration. This seems implausible.} \added{ The H$\alpha$
excess observed by \citet{johnskrull_h_2016} also seems like an
important piece of evidence.  The model with circumstellar optically
thick hydrogen gas near the co-rotation radius readily explains this
excess.  }

\added{Out of the available explanations, we therefore consider the
idea of a clumpy torus of optically thick dust or gas near the
co-rotation radius to be most plausible.  There is theoretical support
for the idea that accretion disks subject to strong magnetic fields
could enter stalled states and become trapped near co-rotation
\citep{dangelo_accretion_2012}.  The concept of an ``evaporating dust
rim'' within which all dust evaporates
\citep{millan-gabet_circumstellar_2007,dullemond_inner_2010} will
likely be an important first-order consideration in physical models of
this scenario.  The additional question of whether the opacity in fact
comes from dust or partially-ionized gas ({\it e.g.}, via gas
continuum, atomic lines, or perhaps TiO molecular opacity) also
remains a topic for future exploration.  }

>>>>

The explanations that \replaced{are not yet}{have not yet been} ruled
out include {\it (i)} transiting clumps of \added{dust or }gas at the
Keplerian co-rotation radius; {\it (ii)} transits of enshrouded
protoplanets; {\it (iii)} occultations of \replaced{starspots}{hot and
cold spots on the stellar surface} by an optically thick disk.  The
first and last explanations have added appeal because they are
flexible enough to explain not only the transient and persistent-dip
M-dwarfs, but also the ``scallop shell'' M-dwarfs
\citep{stauffer_orbiting_2017}.  \replaced{Despite this appeal, the
possibility of distinct mechanisms explaining these distinct
variability classes remains open.}{ In the spot occultation scenario
\citep{zhan_complex_2019}, the ``blobs'' that provide the necessary
brightness contrast are on the surface of the star, rather than at the
co-rotation radius. The observing geometry would require an inner ring
to be projected into a narrow band in front of the stellar disk, in
order to produce variations with sufficiently short timescales.  The
outer disk would also need to either not be present, or to be
misaligned enough and insufficiently flared, that it did not fall
along the line of sight.  This geometric requirement seems too
contrived to occur in even $\lesssim$1\% of young M dwarf systems. }

\replaced{The evolution of \ptfo\ over}{Observations of \ptfo\ from}
the past decade could offer additional hints.  \deleted{Specifically,
}\ptfo's transition between having none, one, and multiple dips per
cycle seems important \added{(Figure~\ref{fig:o_minus_c})}.
\deleted{The dips would then need to be caused by material that was
somehow disrupted from the planet, but somehow remained co-orbital for
an extended duration. This seems implausible.} \added{ The H$\alpha$
excess observed by \citet{johnskrull_h_2016} also seems like an
important piece of evidence.  The model with circumstellar optically
thick hydrogen gas near the co-rotation radius readily explains this
excess.  }

\added{Out of the available explanations, we therefore consider the
idea of a clumpy torus of optically thick dust or gas near the
co-rotation radius to be most plausible.  There is theoretical support
for the idea that accretion disks subject to strong magnetic fields
could enter stalled states and become trapped near co-rotation
\citep{dangelo_accretion_2012}.  The concept of an ``evaporating dust
rim'' within which all dust evaporates
\citep{millan-gabet_circumstellar_2007,dullemond_inner_2010} will
likely also be an important first-order consideration in physical models of
this scenario.  The additional question of whether the opacity in fact
comes from dust or partially-ionized gas ({\it e.g.}, via gas
continuum, atomic lines, or perhaps TiO molecular opacity) also
remains a topic for future exploration.  }
